---
title: "Cast and Join"
format: html
editor: visual
---

# 11/20 Casting and Joining

setup per usual

```{r}
rm (list =ls())
library(ggfortify)
library(tidyverse)
library(here)
```

read in data to use

```{r}
cams <- read.csv(here("data", "site_locations.csv"))
species <- read.csv(here("data", "species_per_site.csv"))
```

pivot_wider() function to change the format so we have a column for each date

```{r}
sp_wide<-species%>%
  pivot_wider(
    names_from=Date,
    values_from=Num_species
  )
View(sp_wide)
```

pivot_longer() function to take the y data frame and mush columns into one long column

```{r}
sp_long<-sp_wide%>%
  pivot_longer(
    cols=c(`1-Jul-24`:`15-Jul-24`),
    names_to="Date", 
    values_to="Num_species"
  )
View(sp_long)
```

## Joining Dataframes

if two dataframes have at least one of the same variable, we can match them and create one big dataframe

getting the latitude and longitude data from cams and add it to species using join() function

```{r}
cams_species<-full_join(species,cams) #df1 gets added to from df2, full_join(df1,df2), use "by=join_by(variable)" to assign the shared column
View(cams_species)
```
