---
title: "Practice with t-tests"
output: html_document
date: "2022-10-27"
---

## Reviewing the t-test

Let's conduct a 2-sample t-test to compare body fat percentage between men and women.

The data set presents the body fat percentage for a sample of men and women who all work out at the same gym. In general, men are leaner than women, with a body fat percentage centering around 15% whereas for women it tends to center around 20%. Let’s see if the same is true of these data.

```{r}
rm (list=ls())
library(tidyverse)
library(here)
```

```{r}
fat <- read.csv(here("data", "body_fat_percentage.csv"))
```

## Instructions

Use the information you learned from reading chapter 5 and doing homework 21 to conduct a t-test comparing mean body fat percentage between the male and female groups. But don’t just rush to the test – follow the recommended analysis workflow:

1.  Plot the data and look at expected relationships

    ```{r}
    fat$gender <- as.factor (fat$gender)
    ```

```{r}

ggplot(fat,aes(x=body_fat_percent))+
  geom_histogram(binwidth=5)+
  facet_wrap(~gender,ncol=1) 
  theme_bw()
```

```{r}
ggplot (fat, aes (x=gender,y=body_fat_percent))+
  geom_boxplot() + # makes a boxplot
  geom_point(size=3,colour='darkred',alpha=0.5)+ # not totally sure what overlay point data at the top means, so I'm just going to increase the size of the points
  xlab("gender") + #xaxis label
  ylab("fat percentage")+ #yaxis label
  theme_bw()
```

2.  Calculate the mean and standard deviation for the data and examine the resulting output.

```{r}
fat %>%
  group_by(gender)%>%
  summarise(
    mean_bf<-mean(body_fat_percent, na.rm=T),
    sd_bf<-sd (body_fat_percent, na.rm=T)
  )

```

1.  

    3.  Discuss with your partner:

    -   Review in the book: What are the assumptions about the data appropriate for a t-test? When you examine the plots and summary data, do they appear to meet those assumptions?

    look for normallish distributed, independent data, equal variance

    when eyeballing the data, we can very roughly assume a t

    -   Based on examination of the plots and summary data, do you expect that there is a statistically significant difference in mean body fat percentage between males and females in this sample?

    yes we expect a statistically significant difference

    4.  Conduct a t-test and then interpret the results. What do they tell you about the null vs. alternative hypotheses?

    looking at a difference between two means, so we're running a t-test

    ```{r}
    t.test(body_fat_percent ~ gender, data=fat)
    ```

p is very low, we reject the null hypothesis (t = 2.8632, df = 20.957, p-value = 0.00932)

it is very likely that men have significantly less body fat than women
