---
title: "Practice with t-tests"
output: html_document
date: "2022-10-27"
---

## Reviewing the t-test

Let's conduct a 2-sample t-test to compare body fat percentage between men and women.

The data set presents the body fat percentage for a sample of men and women who all work out at the same gym. In general, men are leaner than women, with a body fat percentage centering around 15% whereas for women it tends to center around 20%. Let’s see if the same is true of these data.

```{r}
rm (list=ls())
library(tidyverse)
library(here)
```

```{r}
fat <- read.csv(here("data", "body_fat_percentage.csv"))
```

## Instructions

Use the information you learned from reading chapter 5 and doing homework 21 to conduct a t-test comparing mean body fat percentage between the male and female groups. But don’t just rush to the test – follow the recommended analysis workflow:

1.  Plot the data and look at expected relationships

    ```{r}
    fat$gender <- as.factor (fat$gender)
    ```

```{r}

ggplot(fat,aes(x=body_fat_percent))+
  geom_histogram(binwidth=10)+
  facet_wrap(~gender,ncol=1) 
  theme_bw()
```

```{r}
ggplot (fat, aes (x=gender,y=body_fat_percent))+
  geom_boxplot() + # makes a boxplot
  geom_point(size=3,colour='darkred',alpha=0.5)+ # not totally sure what overlay point data at the top means, so I'm just going to increase the size of the points
  xlab("gender") + #xaxis label
  ylab("fat percentage")+ #yaxis label
  theme_bw()
```

2.  Calculate the mean and standard deviation for the data and examine the resulting output.

```{r}
fat %>%
  group_by(gender)%>%
  summarise(
    mean(body_fat_percent),
    sd (body_fat_percent)
  
```

1.  

    3.  Discuss with your partner:

    -   Review in the book: What are the assumptions about the data appropriate for a t-test? When you examine the plots and summary data, do they appear to meet those assumptions?

    normallish distributed, independent data,

    -   Based on examination of the plots and summary data, do you expect that there is a statistically significant difference in mean body fat percentage between males and females in this sample?

    yes we expect a statistically significant difference

2.  Conduct a t-test and then interpret the results. What do they tell you about the null vs. alternative hypotheses?

    looking at a difference between two means, so we're running a t-test

    ```{r}
    t.test(body_fat_percent ~ gender, data=fat)
    ```

p is very low, we reject the null
