---
title: "Oct-9-2024"
author: "ML"
format: html
editor: visual
---

10/09/2024

## Exercise 1

Warmup

set things up and read in tomato data

```{r}
rm(list = ls()) #clears environment
# establish packages
library (tidyverse)
library (here)
```

```{r}
tomatoes <- read.csv (here("data", "tomato_data.csv")) 
```

Write a line of code that returns the 3rd, 7th, and 9th rows of data.

```{r}
tomatoes %>%
  slice (c(3,7,9)) %>%
    View()
```

Write a line of code that finds all of the plants from the fertilizer treatment with avg_tomato_weight_g \> or equal to 145 g.

```{r}
tomatoes %>%
  filter (Avg_tomato_weight_g>=145)

```

save changes to the dataset to a new file in an output folder

```{r}
dir.create(here("outputs"))
```

now save the file

```{r}
write.csv(tomatoes,here("outputs", "tomatoes_fixed.csv"),row.names=FALSE) # prevents an extra column with row numbers to the file
tomatoes2<-read.csv(here("outputs", "tomatoes_fixed.csv"))
```

## NoCoLead Data

```{r}
# read in data
lead <- read.csv (here ("data", "NoCoLead.csv"))
```

find avg num of tests given in the data set

```{r}
lead %>% 
  summarise(
      meantests=mean(num_tests)
      ) %>%
        View()
```

find and save the avg num of tests per year

```{r}
leadsummary <-lead %>%
 group_by(Year)%>%
   summarise(
    meantestsyearly=mean(num_tests)
      )
```

find the

```{r}

```
