---
title: "21 Oct 2024 - review"
format: html
editor: visual
---

Download the dataset called "birds_and_mammals.csv" from today's module and put it in your Data folder.

Set up your workspace and read in the .csv. Name it "critters"

```{r}
# clear environment
rm (list = ls())
library (tidyverse)
library (here)

# get dataset
critters <- read.csv (here("data", "birds_and_mammals.csv"))
```

Filter just for mammals in the order Carnivora and make a new dataset called "carnivores"

```{r}

  carnivores <- critters %>% 
    filter (Order == "Carnivora")
    
```

create a histogram of maximum longevity for carnivores.

```{r}
ggplot(carnivores, aes(max_longevity_years))+ 
  geom_histogram()
```

Now make a scatterplot comparing max_longevity_years vs. adult_wt_g

```{r}
ggplot (carnivores, aes(x=max_longevity_years, y=adult_wt_g))+
  geom_point()
```

Use code (rather than scrolling the data) to determine which species it is that has the highest maximum longevity?

```{r}
critters %>% 
  group_by (Species) %>%
    arrange(desc(max_longevity_years))
# aedium is mysticetus
```

Use code to determine the two species between 2.5e+05 and 3e+05 in adult body mass.

```{r}
carnivores %>% 
  group_by (Species) %>%
    filter (adult_wt_g >=250000, adult_wt_g <= 300000)
```

Now make a boxplot of max longevity as a function of family. Set the y-axis label to "maximum longevity, years". Make the points red and somewhat transparent.

```{r}

```
