---
title: "October-7-2024"
author: "ML"
format: html
editor: visual
---

Oct 7 In Class Excercise 1

```{r}
rm(list = ls())
library (tidyverse)
library (here)
```

## data types for Tomato

Plant number - nominal Treatment - binary/nominal Num flowers - discrete Tomato yield fruits - discrete Avg Tomato Weight g - ration

```{r}
tomatoes <- read.csv (here ("data", "tomato_data.csv"))
tomatoes$Plant_number<- as.factor (tomatoes$Plant_number)
tomatoes$Treatment<- as.factor (tomatoes$Treatment)
levels (tomatoes$Treatment)

fixTreatment <- which (tomatoes$Treatment == ("control")) 
  tomatoes$Treatment [fixTreatment] <- "Control"
fixTreatment <- which (tomatoes$Treatment == ("fertilizer")) 
  tomatoes$Treatment [fixTreatment] <- "Fertilizer"

  tomatoes$Treatment <- factor (tomatoes$Treatment)
```

## summary ()

```{r}
test <- c(2,4,6,8,10,12,14,16,18,20)
summary (test)
```

manually calculate summary values of this data set graze \<- c(90,94,96,90,84,86) 84,86,90,90,94,96

min =84 max =96 mean (96+84)/2 =90 median =90 1Q (96-84)(0.25)+84 =87 3Q (96-84)(0.75)+84 =93

```{r}
grades <- c(90,94,96,90,84,86)
summary (grades)
```

yayyy! we got the same values

```{r}
quantile(grades, probs = c(0, 0.25, 0.5, 0.75, 1))
```

this gives the quantile values at the portions we want, we gave it quarters but could do something else

```{r}
quantile(grades, probs = c(0, 0.33,0.66, 1))
```

## back to Tomato Data with dplyr package
```{r}
# Write a line of code that returns just the column with number of flowers
select (tomatoes)
```

Write a line of code that returns the 3rd through 8th rows of data
Write a line of code that returns only the data that were from the control treatment in the experiment
Write a line of code that returns plants from the fertilized treatment that had at least 27 flowers
Write code to determine the mean number of flowers in the control treatment plants
Write code to determine the mean weight of tomatoes from the fertilizer treatment plants
Use piping to do the following:
Create a new variable called "big_toms" that consists of a data frame consisting of a single column of only the plant numbers from plants that produced fruit with a mean weight of 150 g or more
Use group_by and summarize to determine the mean number of flower produced in the control and treatment groups
Use group_by and summarize to determine the mean and standard deviation in mean fruit size (grams) between control and fertilized plants and save the result in a dataframe called TomatoSizes
